Andres Karjus

### 2021-04-21 index.Rmd
news

<div style="text-align: justify; padding-left:1px"> 
I am a senior research fellow at the [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, working on **language and culture dynamics** using large corpora, machine learning, but also human experiments. I recently defended my PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/) on lexical competition and communicative needs in language, supervised by [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/).
I am also engaged in teaching **stats and data visualization** for the humanities and social sciences (see the tab above). On the right is the (of course slightly outdated) photo obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf) | [1-page non-technical summary](https://andreskarjus.github.io/thesis/english_summary.pdf) | [eestikeelne lühikokkuvõte](https://andreskarjus.github.io/thesis/doktoritoo_ylevaade_AK.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | in review | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [journal link](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

@@

## Teaching, consulting & R workshops {#workshops}

If you're here about materials for an upcoming dataviz workshop: [click here](https://andreskarjus.github.io/artofthefigure/) to go to the dedicated workshops page.

@@

- I have been developing teaching materials on data visualization using R, which I have used in over a dozen workshops for humanities and social science audiences since 2017, collectively under the name of [_aRt of the Figure_ (click here for more info)](https://andreskarjus.github.io/artofthefigure/). Some of these have been standalone events, some have been part of conferences, summer schools or academic retreats. All my materials are [open source](https://andreskarjus.github.io/artofthefigure/).

@@

- My postdoc does not include teaching activities, but I do occasional guest lectures; most recently for the Data Science and Digital Humanities programme at the University of Tartu.

@@

<div style="text-align: center; align:center"> 
<IMG SRC="round.jpg"> 
<br>
<br>
<br>
Andres Karjus, senior research fellow, CUDAN lab, Tallinn University<br>
&nbsp;&nbsp;&nbsp;<a href="mailto:andres.karjus --at-- tlu.ee">email</a>  |  <a href="https://twitter.com/AndresKarjus">twitter</a>
<br>
</div>

>>>>

<div style="text-align: justify; padding-left:1px"> 
I am a research fellow at the [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, working on **language and culture dynamics** using large corpora, machine learning, but also human experiments. I recently defended my PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/) on lexical competition and communicative needs in language, supervised by [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/).
I am also engaged in teaching **stats and data visualization** for the humanities and social sciences (see the tab above). On the right is the (of course slightly outdated) photo obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Andres Karjus, Richard A. Blythe, Simon Kirby, Tianyu Wang, Kenny Smith. 2021. Conceptual similarity and communicative need shape colexification" | [preprint](arxiv.org/abs/2103.11024) | [code and data](https://github.com/andreskarjus/colexification_experiment)
* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf) | [1-page non-technical summary](https://andreskarjus.github.io/thesis/english_summary.pdf) | [eestikeelne lühikokkuvõte](https://andreskarjus.github.io/thesis/doktoritoo_ylevaade_AK.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | in review | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [journal link](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

@@

## R workshops {#workshops}

If you're here about materials for an upcoming R or dataviz workshop: [click here](https://andreskarjus.github.io/artofthefigure/) to go to the dedicated workshops page.

@@

- I have been developing teaching materials on data visualization using R, which I have used in over a dozen workshops for humanities and social science audiences since 2017, collectively under the title of [_aRt of the Figure_ (click here for more info)](https://andreskarjus.github.io/artofthefigure/). Some of these have been standalone events, some have been part of conferences, summer schools or academic retreats. All my materials are [open source](https://andreskarjus.github.io/artofthefigure/).

@@

- My postdoc does not include teaching activities, but I do occasional guest lectures; most recently for the Data Science and Digital Humanities programme at the University of Tartu and for the Cultural Data Analytics I course at Tallinn University.

@@

<div style="text-align: center; align:center"> 
<IMG SRC="round.jpg"> 
<br>
<br>
<br>
Andres Karjus, research fellow, CUDAN lab, Tallinn University<br>
&nbsp;&nbsp;&nbsp;<a href="mailto:andres.karjus --at-- tlu.ee">email</a>  |  <a href="https://twitter.com/AndresKarjus">twitter</a>
<br>
</div>
